Dietary niche relationships between core bumble bee species in the Berchtesgadener Alps

### 2020-11-26 analyses/niche_relations_v4.Rmd
polished tree plots

## Define functions
```{r, echo = FALSE}
# Convert PCoA to tidy array ("fortify")
pcoa_fort <- function(x){
  fortify(x, axes = 1:2) %>% 
  mutate(Score = case_when( # rename default scores to match my data
    Score == "species" ~ "FL.sp",
    Score == "sites" ~ "BB.sp"))
}

@@

  ggplot(filter(x, (abs(MDS1) > min_score | abs(MDS2) > min_score) | (Score == "BB.sp"))) + # no threshold for BB spp
    geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), # draw polygons around T classes
                 color = "gray70", fill = NA, size = 1.25, key_glyph = "abline") +
    geom_point(aes(MDS1, MDS2, color = Score), size = 3, alpha = 0.75) + # add points
    ggrepel::geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect") + # add labels
    coord_fixed() + # Gavin Simpson says this is **really** important
    theme_classic() 
}

@@

## Extract and name plots
niche_oben_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Oben/Early/Tax")
niche_oben_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Oben/Mid/Tax")
niche_oben_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Oben/Late/Tax")

niche_mitte_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Mitte/Early/Tax")
niche_mitte_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Mitte/Mid/Tax")
niche_mitte_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Mitte/Late/Tax")

niche_unten_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Unten/Early/Tax")
niche_unten_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Unten/Mid/Tax")
niche_unten_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Unten/Late/Tax")

@@

# Relative diet breadth ~ proportional abundance
ggplot(diet_breadth.tax, aes(prop, normalised.degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "lm")
```
### Results
In a nutshell, proportional diet breadth is extremely tightly correlated with proportional anbundance. The story here is that LT species are more abundant than ST species at lower elevations, and ST species are more abundant than LT species at higher elevations. This relationship is strong and basically linear, with the relative abundances "crossing" at

@@

## 1.6 Tree plots
This visualization deserves to be pulled out as a stand-alone chunk. It's so good.
```{r}
### Visitation by BB species, parsed by elev.class and period
visitation_abund.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
  group_by(elev.class) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
  mutate(k.type.ss = factor(k.type.ss),
         bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
  group_by(period, elev.class, plant.genus, k.type.ss, plant.family, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(elev.class, period, bb.sp) %>%
  mutate(prop = abund/sum(abund)) %>%
  arrange(elev.class, period, bb.sp, -prop)

@@

# Early
ggplot(filter(visitation_abund.sp, period == "early"), 
       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
  geom_treemap(color = "black") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols) + 
  facet_grid(fct_rev(elev.class)~bb.sp)

# Mid
ggplot(filter(visitation_abund.sp, period == "mid"), 
       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
  geom_treemap(color = "black") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols) + 
  facet_grid(fct_rev(elev.class)~bb.sp)

# Late
ggplot(filter(visitation_abund.sp, period == "late"), 
       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
  geom_treemap(color = "black") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols) + 
  facet_grid(fct_rev(elev.class)~bb.sp)

>>>>

## Define functions
```{r, echo = FALSE}
# Convert PCoA to tidy array ("fortify")
pcoa_fort <- function(x){
  fortify(x, axes = 1:2) %>% 
  mutate(Score = case_when( # rename default scores to match my data
    Score == "species" ~ "FL sp.",
    Score == "sites" ~ "BB sp."))
}

@@

  ggplot(filter(x, (abs(MDS1) > min_score | abs(MDS2) > min_score) | (Score == "BB.sp"))) + # no threshold for BB spp
    geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), # draw polygons around T classes
                 color = "gray70", fill = NA, size = 1.25, key_glyph = "abline") +
    geom_point(aes(MDS1, MDS2, color = Score), size = 3, alpha = 0.75) + # add points
    ggrepel::geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect", size = 3) + # add labels
    coord_fixed() + # Gavin Simpson says this is **really** important
    theme_classic() +
    theme(legend.title=element_blank())
}

@@

## Extract and name plots
niche_oben_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "early")$pcoa.plot[[1]] 
niche_oben_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "mid")$pcoa.plot[[1]] 
niche_oben_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "late")$pcoa.plot[[1]]

niche_mitte_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "early")$pcoa.plot[[1]]
niche_mitte_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "mid")$pcoa.plot[[1]] 
niche_mitte_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "late")$pcoa.plot[[1]]

niche_unten_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "early")$pcoa.plot[[1]]
niche_unten_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "mid")$pcoa.plot[[1]]
niche_unten_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "late")$pcoa.plot[[1]]

@@

# Relative diet breadth ~ proportional abundance
ggplot(diet_breadth.tax, aes(prop, normalised.degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "lm")
```
### Results
In a nutshell, proportional diet breadth is extremely tightly correlated with proportional anbundance. The story here is that LT species are more abundant than ST species at lower elevations, and ST species are more abundant than LT species at higher elevations. This relationship is strong and basically linear, with the relative abundances crossing at mid-elevation.

@@

## 1.6 Tree plots
This visualization deserves to be pulled out as a stand-alone chunk. It's so good.
```{r}
### Visitation by BB species, parsed by elev.class and period
visitation_abund.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
  group_by(elev.class) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
  mutate(k.type.ss = factor(k.type.ss),
         bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
  group_by(period, elev.class, plant.sp.abb, plant.genus, k.type.ss, plant.family, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(elev.class, period, bb.sp) %>%
  mutate(prop = abund/sum(abund)) %>%
  arrange(elev.class, period, bb.sp, -prop)

@@

# Early
tp.early <- ggplot(filter(visitation_abund.sp, period == "early"), 
                aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
  geom_treemap(color = "white") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols) + 
  labs(fill = "Morphotype") +
  facet_grid(fct_rev(elev.class)~bb.sp)

# Mid
tp.mid <- ggplot(filter(visitation_abund.sp, period == "mid"), 
                 aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
  geom_treemap(color = "white") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols) + 
  labs(fill = "Morphotype") +
  facet_grid(fct_rev(elev.class)~bb.sp)

# Late
tp.late <- ggplot(filter(visitation_abund.sp, period == "late"), 
                  aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
  geom_treemap(color = "white") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols) + 
  labs(fill = "Morphotype") +
  facet_grid(fct_rev(elev.class)~bb.sp)

pdf("./output/tree_ord_early.pdf", width = 14, height = 7)
(tp.early | (niche_oben_early.tax.plot / niche_mitte_early.tax.plot / niche_unten_early.tax.plot) + plot_layout(guides = 'collect')) + plot_layout(widths = c(3, 1))
dev.off()

pdf("./output/tree_ord_mid.pdf", width = 14, height = 7)
(tp.mid | (niche_oben_mid.tax.plot / niche_mitte_mid.tax.plot / niche_unten_mid.tax.plot) + plot_layout(guides = 'collect')) + plot_layout(widths = c(3, 1))
dev.off()

pdf("./output/tree_ord_late.pdf", width = 14, height = 7)
(tp.late | (niche_oben_late.tax.plot / niche_mitte_late.tax.plot / niche_unten_late.tax.plot) + plot_layout(guides = 'collect')) + plot_layout(widths = c(3, 1))
dev.off()
